---
title: "Coeficiente de correlación"
author: "Sofía Fernández"
date: "September 9, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(ggplot2)
theme_set(theme_bw())
```

## Problema

An article by Tuzson et al. (A-9) in Archives of Physical Medicine and Rehabilitation reported the following data on peak knee velocity in walking (measured in degrees per second) at flexion and extension for 18 subjects with cerebral palsy.

1. Prepare a scatter diagram
2. Compute the sample correlation coefficcient
3. Test $H_0:$ \(\rho\) = 0 at the .05 level of significance and state your conclusion
4. Determine the *p* value for the test
5. Construct the 95 percent confidence interval for \(\rho\). 

***
## Datos

Se leyeron los datos de los recursos del libro: 

```{r data, message=FALSE}
knee_vel <- readr::read_csv(file = "data_daniels/ch09_all/EXR_C09_S07_04.csv")
```
```{r data1, echo=F}
knee_vel
```

***
## Gráfica de dispersión

```{r plot}
ggplot(knee_vel, aes(FLEX, EXT)) + 
    geom_point() + 
    geom_smooth(method = "lm", se = F, colour = 'black') +
    geom_vline(aes(xintercept = mean(FLEX)), 
               colour = "gray", linetype = "dashed") + 
    geom_hline(aes(yintercept = mean(EXT)),
               colour = "gray", linetype = "dashed")
```

## Cálculo de correlación
```{r corr}
result <- cor.test(~ EXT + FLEX, data = knee_vel)
```

```{r results, echo=FALSE}
result
```

***
## Resultados

### Coeficiente de correlación
```{r res1, echo=FALSE}
result$estimate
```

### Prueba de hipótesis

$H_0:$ \(\rho\) = 0

```{r res2, echo=FALSE}
result$statistic
```

### Valor *p*
```{r res3, echo=FALSE}
result$p.value
```

### Intervalo de confianza
```{r res4, echo=FALSE}
result$conf.int
```

## Conclusión
Hay una confianza del 95% de que \(\rho\) esté contenida en el intervalo 0.706 a 0.956. Por lo tanto se rechaza la hipótesis nula. 